---
title: "BlindMonkeys"
author: "Andy Pickering"
date: "August 3, 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I was recently reading the book "Black box thinking" by Matthew Syed (<http://www.matthewsyed.co.uk/books/>), and came across an interesting example illustrating the power of natural selection. Apparently it came from the book "The Blind Watchmaker" by Richard Dawkins. I thought it would be fun to try it out in R. The scenario is that a Blind Monkey is pressing keys randomly on a typewriter. What are the odds the monkey will type a specific sentence? Without any selection mechanism, the chances of this happenning are very small and it will take a very long time. But if we add in a selection mechanism (analgous to natural selection), this random typing produces a specific sentence much more quickly, illustrating the enormous power of random variation combined with natural selection.


Below is some R code to simulate this scenario. We start with a specified phrase, randomly generate the same number of letters. Then we keep the letters that are correct (this is our 'natural selection'), and keep repeating this process until we get the exact phrase. Try it out with a word or phrase of your choice! 
```{r}
rm(list=ls())
thePhrase<-c("c","h","e","e","s","e")#"Cheese"
Nchar<-length(thePhrase)
    # generate a random sample
    b<-sample(letters,Nchar)
    # find which characters are correct
    ig<-which(thePhrase==b)
    
    #ig<-0
    n_iter<-1
    while (length(ig)<Nchar){
        # generate a random sample
        b<-sample(letters,Nchar)
        # keep the correct letters
        b[ig]=thePhrase[ig]
        # find which characters are correct
        ig<-which(thePhrase==b)
        print(b)
        n_iter<-n_iter+1
    }
    # How many iterations ('generations') did it take?
    print(n_iter)
    # Print the final result
    print(b)
```






